Mental health issue linked to major crimes

### 2024-09-25 Paper/Paper.qmd
Add files via upload

@@

## 3.1 Results of MHA Apprehensions Data

```{r,fig.pos='H'}
#| label: MHA Apprehensions Data
#| fig-cap: Number of MHA reports in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

```{r,fig.pos='H'}
#| label: MHA Apprehensions Data 2
#| fig-cap: Number of MHA reports in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

@@

## 4.1 Results of Major Crime Indicator Data

```{r,fig.pos='H'}
#| label: Major Crime Indicator Data
#| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

```{r,fig.pos='H'}
#| label: Major Crime Indicator Data 2
#| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

```{r,fig.pos='H'}
#| label: comparison 
#| fig-cap: comparison between two data
#| echo: false
#| warning: false
#| message: false

@@

```{r,fig.pos='H'}
#| label: comparison 2
#| fig-cap: comparison between two data
#| echo: false
#| warning: false
#| message: false

@@

By comparing the two line plots,a positive relationship between the number of MHA reports and the number of Major Crime Indicator, the number of MHA reports seems to influence the number of crime incidents accross different police divisions. In order to confirm there are linear relationship between two factors, a linear model is made to fit in this situation based on two data:

$$
C = 6748.244 + 2.451 \times MHA
$$ in which:

@@

The corresponding graph of linear model is:

```{r,fig.pos='H'}
#| label: Linear model
#| fig-cap: Linear Model
#| echo: false
#| warning: false
#| message: false

@@

The p-values for both intercept and coefficient (MHA Reports) are below the significant level of 0.05, indicating a strong evidence that the relationship between MHA Reports and major crime indicator follows this linear model.

```{r,fig.pos='H'}
#| label: Linear model summary
#| echo: false
#| warning: false
#| message: false

@@

This paper investigated the relationship between the number of Mental Health Act Apprehensions and the major crime indicator, according to Toronto Police Services data retrieved from OpenData Toronto. Results shows that as the number of MHA reports rises, the major crime indicator also increases, based on data from 2014 to 2024. Furthermore, there appears to be a positive linear relationship between the number of MHA reports and the number of major crime indicator, which can be expressed as a linear model: C = 6748.244 + 2.451 \* MHA.

The result of this paper shares the insight of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is 4.19\*10\^-5, which is far below the significance level of 0.05, indicating a strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.[@PublicSafetyCanada]

@@

Significant outliers are evident in both the Mental Health Act (MHA) reports and the Major Crime Indicator data. These outliers may be attributable to the relatively small sample size, given that there are only 20 police divisions in Toronto. The limited number of divisions can lead to greater variability in the data, making it more susceptible to the influence of atypical observations.

```{r,fig.pos='H'}
#| label: Box Plot
#| fig-cap: The box plot of two factors
#| echo: false
#| warning: false
#| message: false

@@

The linear model reveals a significant linear relationship between the number of Mental Health Act (MHA) reports and the major crime indicator. However, the adjusted R-squared value of 0.6388 indicates that approximately 36.12% of the variance in crime rates remains unexplained by the model. This finding implies that, in addition to MHA reports, there are likely other influencing factors contributing to variations in crime rates.

```{r,fig.pos='H'}
#| label: Summary table
#| echo: false
#| warning: false
#| message: false

@@

The residuals in this model are relatively large, leading to a complex residual pattern. This complexity may indicate potential issues with model fit or the presence of unaccounted variability in the data.

```{r,fig.pos='H'}
#| label: Residual Plot
#| fig-cap: The Residual Vs Fitted value plot
#| echo: false
#| warning: false
#| message: false

@@

```{=tex}
\begin{table}[H]
    \centering
    \caption{Linear Model Summary 3}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Statistic} & \textbf{Value} \\ \hline
        Minimum & -10046.5 \\ \hline
        1st Quartile (1Q) & -3273.8 \\ \hline
        Median & -423.9 \\ \hline
        3rd Quartile (3Q) & 4260.7 \\ \hline
        Maximum & 6344.1 \\ \hline
        \textbf{Coefficients} &  \\ \hline
        (Intercept) & 6748.2438 \\ \hline
        Std. Error (Intercept) & 2970.8029 \\ \hline
        t value (Intercept) & 2.272 \\ \hline
        Pr(>|t|) (Intercept) & 0.0373 * \\ \hline
        num\_mha\_reports & 2.4509 \\ \hline
        Std. Error (num\_mha\_reports) & 0.4397 \\ \hline
        t value (num\_mha\_reports) & 5.574 \\ \hline
        Pr(>|t|) (num\_mha\_reports) & 4.19e-05 *** \\ \hline
        \textbf{Significance Codes} & 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 \\ \hline
        \textbf{Residual Standard Error} & 4835 on 16 degrees of freedom \\ \hline
        \textbf{Multiple R-squared} & 0.6601 \\ \hline
        \textbf{Adjusted R-squared} & 0.6388 \\ \hline
        \textbf{F-statistic} & 31.07 on 1 and 16 DF \\ \hline
        \textbf{p-value of F-statistic} & 4.194e-05 \\ \hline
    \end{tabular}
\end{table}
```

```{=tex}
\begin{table}[H]
    \centering
    \caption{Cleaned Data}
    \begin{tabular}{|l|r|r|}
        \hline
        \textbf{Division} & \textbf{Number of Crimes} & \textbf{Number of MHA Reports} \\ \hline
        D11 & 15894 & 5435 \\ \hline
        D12 & 17151 & 5137 \\ \hline
        D13 & 14180 & 4226 \\ \hline
        D14 & 27590 & 12603 \\ \hline
        D22 & 26578 & 6302 \\ \hline
        D23 & 25782 & 5769 \\ \hline
        D31 & 29099 & 6531 \\ \hline
        D32 & 31273 & 7528 \\ \hline
        D33 & 17836 & 5516 \\ \hline
        D41 & 27591 & 6752 \\ \hline
        D42 & 24037 & 5357 \\ \hline
        D43 & 27509 & 7069 \\ \hline
        D51 & 32666 & 10233 \\ \hline
        D52 & 23487 & 5916 \\ \hline
        D53 & 21199 & 6584 \\ \hline
        D54 & 6347 & 1965 \\ \hline
        D55 & 24524 & 8636 \\ \hline
        NSA & 3992 & 755 \\ \hline
    \end{tabular}
\end{table}
```

\vspace{2em}
\newpage

# 9.Reference

>>>>

The bar plot (Figure @fig-MHA) and the line plot (Figure @fig-MHA-2) show the number of MHA Apprehensions in each police division and its trend.

@@

## 3.1 Overview of MHA Apprehensions Data

```{r,fig.pos='H'}
#| label: fig-MHA
#| fig-cap: Number of MHA reports in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

```{r,fig.pos='H'}
#| label: fig-MHA-2
#| fig-cap: Number of MHA reports in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

The bar plot (Figure @fig-MCI) and the line plot (Figure @fig-MCI-2) show the number of Major Crime Indicators in each police division and its trend.

@@

## 4.1 Overview of Major Crime Indicator Data

```{r,fig.pos='H'}
#| label: fig-MCI
#| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

```{r,fig.pos='H'}
#| label: fig-MCI-2
#| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

@@

## 5.1 Comparing two datasets
```{r,fig.pos='H'}
#| label: fig-compare
#| fig-cap: comparison between two data
#| echo: false
#| warning: false
#| message: false

@@

```{r,fig.pos='H'}
#| label: fig-compare-2
#| fig-cap: comparison between two data
#| echo: false
#| warning: false
#| message: false

@@

By comparing the two bar plots (Figure @fig-compare) and line plots (Figure @fig-compare-2),there seems to be a positive relationship between the number of MHA reports and the number of Major Crime Indicator. To check whether they are linearly related, a test for linear relationship is necessary.

## 5.2 Introducing linear model
In order to confirm there are linear relationship between two factors, a linear model is made to fit in this situation based on two data sets:

$$
C = 6748.244 + 2.451 \times MHA 
$$

in which:

@@

The corresponding graph of linear model is @fig-linear-model

```{r,fig.pos='H'}
#| label: fig-linear-model
#| fig-cap: Linear Model
#| echo: false
#| warning: false
#| message: false

@@

The p-values for both intercept and coefficient (MHA Reports) are below the significant level of 0.05 (Figure @fig-Summary), indicating a strong evidence that the relationship between MHA Reports and major crime indicator follows this linear model.

```{r,fig.pos='H'}
#| label: fig-Summary
#| fig-cap: Linear Model Summary 1
#| echo: false
#| warning: false
#| message: false

@@

This paper investigated the relationship between the number of Mental Health Act Apprehensions and the major crime indicator, according to Toronto Police Services data retrieved from OpenData Toronto. Results shows that as the number of MHA reports rises, the major crime indicator also increases, based on data from 2014 to 2024. Furthermore, there appears to be a positive linear relationship between the number of MHA reports and the number of major crime indicator, which can be expressed as a linear model $C = 6748.244 + 2.451 \times MHA$.

The result of this paper shares the insight of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is $4.19 \times 10^{-5}$, which is far below the significance level of 0.05, indicating a strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.[@PublicSafetyCanada]

@@

Significant outliers are evident in both the Mental Health Act (MHA) reports and the Major Crime Indicator data (Figure @fig-Box-Plot). These outliers may be attributable to the relatively small sample size, given that there are only 20 police divisions in Toronto. The limited number of divisions can lead to greater variability in the data, making it more susceptible to the influence of atypical observations.

```{r,fig.pos='H'}
#| label: fig-Box-Plot
#| fig-cap: The box plot of two factors
#| echo: false
#| warning: false
#| message: false

@@

The linear model reveals a significant linear relationship between the number of Mental Health Act (MHA) reports and the major crime indicator. However, the adjusted R-squared value of 0.6388 indicates that approximately 36.12% of the variance in crime rates remains unexplained by the model (Figure @fig-Summary-2). This finding implies that, in addition to MHA reports, there are likely other influencing factors contributing to variations in crime rates.

```{r,fig.pos='H'}
#| label: fig-Summary-2
#| fig-cap: Linear Model Summary 2
#| echo: false
#| warning: false
#| message: false

@@

The residuals in this model are relatively large, leading to a complex residual pattern (Figure @fig-Residual-Plot). This complexity may indicate potential issues with model fit or the presence of unaccounted variability in the data.

```{r,fig.pos='H'}
#| label: fig-Residual-Plot
#| fig-cap: The Residual Vs Fitted value plot
#| echo: false
#| warning: false
#| message: false

@@

```{=tex}
\begin{table}[H]
    \centering
    \caption{Linear Model Summary 3}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Statistic} & \textbf{Value} \\ \hline
        Minimum & -10046.5 \\ \hline
        1st Quartile (1Q) & -3273.8 \\ \hline
        Median & -423.9 \\ \hline
        3rd Quartile (3Q) & 4260.7 \\ \hline
        Maximum & 6344.1 \\ \hline
        \textbf{Coefficients} &  \\ \hline
        (Intercept) & 6748.2438 \\ \hline
        Std. Error (Intercept) & 2970.8029 \\ \hline
        t value (Intercept) & 2.272 \\ \hline
        Pr(>|t|) (Intercept) & 0.0373 * \\ \hline
        num\_mha\_reports & 2.4509 \\ \hline
        Std. Error (num\_mha\_reports) & 0.4397 \\ \hline
        t value (num\_mha\_reports) & 5.574 \\ \hline
        Pr(>|t|) (num\_mha\_reports) & 4.19e-05 *** \\ \hline
        \textbf{Significance Codes} & 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 \\ \hline
        \textbf{Residual Standard Error} & 4835 on 16 degrees of freedom \\ \hline
        \textbf{Multiple R-squared} & 0.6601 \\ \hline
        \textbf{Adjusted R-squared} & 0.6388 \\ \hline
        \textbf{F-statistic} & 31.07 on 1 and 16 DF \\ \hline
        \textbf{p-value of F-statistic} & 4.194e-05 \\ \hline
    \end{tabular}
\end{table}
```
```{=tex}
\begin{table}[H]
    \centering
    \caption{Cleaned Data}
    \begin{tabular}{|l|r|r|}
        \hline
        \textbf{Division} & \textbf{Number of Crimes} & \textbf{Number of MHA Reports} \\ \hline
        D11 & 15894 & 5435 \\ \hline
        D12 & 17151 & 5137 \\ \hline
        D13 & 14180 & 4226 \\ \hline
        D14 & 27590 & 12603 \\ \hline
        D22 & 26578 & 6302 \\ \hline
        D23 & 25782 & 5769 \\ \hline
        D31 & 29099 & 6531 \\ \hline
        D32 & 31273 & 7528 \\ \hline
        D33 & 17836 & 5516 \\ \hline
        D41 & 27591 & 6752 \\ \hline
        D42 & 24037 & 5357 \\ \hline
        D43 & 27509 & 7069 \\ \hline
        D51 & 32666 & 10233 \\ \hline
        D52 & 23487 & 5916 \\ \hline
        D53 & 21199 & 6584 \\ \hline
        D54 & 6347 & 1965 \\ \hline
        D55 & 24524 & 8636 \\ \hline
        NSA & 3992 & 755 \\ \hline
    \end{tabular}
\end{table}
```
```{=tex}
\vspace{2em}
\newpage
```
# 9.Reference
